## Exploration of Financial Contributions for the 2016 Presidential Campaign 

#### by Sim Virdi

***

## Abstract

This report explores a dataset containing approximately 437000 Financial Contributions to the 2016 Presidential Campaigns.

I retrieved US Zip Codes from this [website](https://www.gaslampmedia.com/download-zip-code-latitude-longitude-city-state-county-csv/).

## Dataset

```{r message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
library(gridExtra)
library(RColorBrewer)
library(reshape2)
library(dplyr)
library(tidyr)
library(lubridate) # Might be needed for dates, See problem set 1
library(GGally)
library(scales)
library(memisc)
library(maps) # Needed for plotting NY State

```

```{r echo=FALSE, Load_the_Data}
# Load the Data
setwd("/Users/Sim/Desktop/Nanodegrees/Data Analyst/Task 4/Data Analysis with R/Assignment/")
data <- read.csv("NY_contributions.csv", row.names = NULL)
zips <- read.csv("zip_codes_states.csv")
parties <- read.csv("cand_gend_party.csv")

# Remove incorrect column from the dataset
colnames(data) <- c(colnames(data)[-1],"x")
data$x <- NULL

# Renaming columns in the data to be less verbose
data <- rename(data, c("contbr_city" = "city",
                       "contbr_st" = "st",
                       "contbr_zip" = "zip",
                       "contbr_employer" = "employer",
                       "contbr_occupation" = "occupation",
                       "contb_receipt_amt" = "amount",
                       "contb_receipt_dt" = "date"))

zips <- rename(zips, "zip_code" = "zip")

# Zipcodes need cleaning because majority are postal codes, only take first 5 characters
data$zip <- substr(data$zip, 0, 5)
data$zip <- ifelse(nchar(data$zip) < 4, 
                   NA,
                   data$zip)
#subset(data$zip, nchar(data$zip) < 4)
#unique(data$zip)

mdata <- merge(data, zips[, c("zip","latitude","longitude", "state")], by = "zip")
mdata <- merge(mdata, parties, by = "cand_nm")

# removing zip codes that aren't from NY state
mdata <- subset(mdata, mdata$state == "NY")

# Only interested in general and primary election from 2016
mdata <- subset(mdata, mdata$election_tp == "P2016" | mdata$election_tp == "G2016")
```

***

## Univariate Plots Section

```{r echo=FALSE, Univariate_Plots}
dim(mdata)
str(mdata)
summary(mdata)
```

### Distribution in State

In the plot below, you can visualise the distribution of donations from NY State, however it is uneasy to see both the state and city boundaries. 
```{r echo=FALSE, warning=FALSE, Long_and_Lat}
# Plotting the long and lat of the donations
ggplot(aes(x = longitude, y = latitude), 
       data = subset(mdata, 
                     !is.na(mdata$zip))) +
  geom_point()
```

To make the boundaries clearer, I used the following [tutorial](https://uchicagoconsulting.wordpress.com/tag/r-ggplot2-maps-visualization/). 

From the plot, you can see that the largest cities have the highest frequency of contributions.
```{r echo=FALSE, warning=FALSE, Map_of_New_York}
# Choosing new york state
all_states <- map_data("state")
newyork <- subset(all_states, region %in% "new york")

# Drawing silouhette of NY with contributor data points
ggplot() +
  geom_polygon(data = newyork,
               aes(x = long, y = lat, group = group)) +
  geom_jitter(aes(x = longitude, y = latitude, group = NULL), 
             data = subset(mdata, 
                     !is.na(mdata$zip)),
             alpha = 1/50,
             color = "coral1") +
  geom_point(data = subset(us.cities,
                           us.cities$country == "NY" & us.cities$pop > 100000),
             aes(x = long, y = lat, size = pop),
             colour = "green") +
  geom_text(data = subset(us.cities,
                          us.cities$country == "NY" & us.cities$pop > 100000),
            aes(x = long, y = lat, label = name),
            colour = "gold",
            fontface="bold")
```


### Amount Contributed

The next plot is used to visualise the amount of money being contributed. 

From the plot, you can see there are negative amounts being displayed due to refunds (we will explore them in further detail later).

```{r echo=FALSE, warning=FALSE, Amounts_with_refunds}
# Plot with refunds
qplot(x = amount, data = mdata) +
  scale_x_continuous(limits = c(-400, 400), breaks = seq(-400,400,50))

dim(subset(mdata, mdata$amount < 0))

# Summary with refunds.
summary(mdata$amount)

```

After removing the refunds, we see a positive skew in the distribution, with most of the money contributed between $15-$100. The median contribution is 27 with an IQR of 85. The mean is quite heavily skewed (due to v. large contributions).

```{r echo=FALSE, warning=FALSE, Amounts_without_refunds}
# Plot without refunds
qplot(x = amount, data = subset(mdata, amount >= 0)) +
  scale_x_continuous(limits = c(0, 400), breaks = seq(0,400,20))

# Summary without refunds.
summary(subset(mdata$amount, mdata$amount >= 0))
```

With the amount transformed using log10, we can see a normal distrubiton with a peak at 27.

```{r echo=FALSE, warning=FALSE, Amounts_transformed_with_log_without_refunds}
# Log transformed 
ggplot(aes(amount),
       data = subset(mdata, amount > 0)) +
  geom_histogram() +
  scale_x_continuous(trans = log10_trans(),
                     limits = c(1, 10000),
                     breaks = c(0, 10, 100, 500, 1000, 10000)) 
```

### Contributions by Gender

We are now exploring contribution amounts by gender. 

As can be seen in the histograms, both distributions have a very similar shape (both had a positive skew) with the male category having a higher frequency of contributions at $30 & $50.

From the summaries and boxplots respectively, we can see that Females have a larger mean and median contribution amount. Females also have anomalies at higher amounts.

From the frequency polygon, we get a clearer picture about the different amounts contributed to gender with females only having higher contributions at $20 and $70.
```{r echo=FALSE, warning=FALSE, Contributions_by_Gender}
qplot(x = amount, data = subset(mdata, amount >= 0)) +
  scale_x_continuous(limits = c(0, 200), breaks = seq(0,200,20)) +
  facet_wrap(~gender)

table(mdata$gender)
by(mdata$amount, mdata$gender, summary)

qplot(x = gender, y = log10(amount),
      data = subset(mdata, amount > 0),
      geom = 'boxplot')

qplot(x = amount, y = ..count..,
      data = subset(mdata,  amount > 0), 
      xlab = 'Amount Contributed',
      ylab = 'Frequency',
      binwidth = 10, 
      geom = 'freqpoly', 
      color = gender) + 
  scale_x_continuous(lim = c(0, 200), breaks = seq(0, 200, 20)) 

```

### Contributions by Party

We are now exploring contribution amounts by Party. 

From the histogram, we can determine that the plots all have similar shapes with a positive skew. 

From the Frequency polygon, we can see that the Democratic party has significantly more contributions than the Republicans (7.37 times more) and the Green party (1585 times more).

From the summaries and boxplot, we can see that the Democrats have the lowest median and mean (27, 191 respectively) of the 3 parties. The next highest medians are the Republicans with a median of 50 although they have the highest mean of 241.5. The green party has a median of 104.0 and a mean of 201.8 although they received significantly less contributions.

```{r echo=FALSE, warning=FALSE, Contributions_by_Party}
qplot(x = amount, data = subset(mdata, amount >= 0)) +
  scale_x_continuous(limits = c(0, 200), breaks = seq(0,200,20)) +
  facet_wrap(~party, scales = 'free')

qplot(x = amount, y = ..count..,
      data = subset(mdata,  amount > 0), 
      xlab = 'Amount Contributed',
      ylab = 'Frequency',
      binwidth = 10, 
      geom = 'freqpoly', 
      color = party) + 
  scale_x_continuous(lim = c(0, 200), breaks = seq(0, 200, 20)) 

table(mdata$party)
by(mdata$amount, mdata$party, summary)

qplot(x = party, y = log10(amount),
      data = subset(mdata, amount > 0),
      geom = 'boxplot')
```

## Univariate Analysis

### What is the structure of your dataset?
There are 428646 contributions in the dataset with 23 features. 

Observations:

* Most contributions are coming from heavily populated areas such as cities.
* The median contribution is $27 and max contribution is $4904860.51.
* Contributions are spread almost equally across gender.
* Most contributions go to the Democratic Party.

### What is/are the main feature(s) of interest in your dataset?
The main features in the dataset are amount, candidate and party. I'd like to determine which features are the best for predicting the contribution size.

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?
Gender, occupation, candidate, zip and time will likely influence the contribution size.

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?

I log transformed the amount of contribution for right skewed plots so it would be easier to see distributions. The distribution that I thought was most unusual was the contributions by party. I assumed that all parties would have a similar frequency of contributions however it turned out that the Democrats had almost 80% of contributions.

I adjusted column names to make them less verbose and removed rows that had data not related to the Primary or General 2016 elections or not within NY state. 

I inserted columns into the data to hold the longitude/latitude of zipcodes as well as gender and party of candidates. I did this so I could plot the contributions on a map over NY state and so that I could determine how the contributions were split between parties and gender of the candidates.

***

## Bivariate Plots Section
```{r echo=FALSE, Bivariate_Plots}

```

## Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

### What was the strongest relationship you found?


***

## Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}

```

## Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

***

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

***

# Reflection
I wanted to answer the question of whether if you come from a certain town, are you more likely to receive votes from that town.

