## Exploration of Financial Contributions for the 2016 Presidential Campaign 

#### by Sim Virdi

***

## Abstract

## Dataset
### Preperations

```{r message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
library(gridExtra)
library(RColorBrewer)
library(reshape2)
library(dplyr)
library(tidyr)
library(lubridate) # Might be needed for dates, See problem set 1
library(GGally)
library(scales)
library(memisc)
library(maps) # Needed for plotting NY State

```

```{r echo=FALSE, Load_the_Data}
# Load the Data
setwd("/Users/Sim/Desktop/Nanodegrees/Data Analyst/Task 4/Data Analysis with R/Assignment/")
data <- read.csv("NY_contributions.csv", row.names = NULL)
zips <- read.csv("zip_codes_states.csv")

# Remove incorrect column from the dataset
colnames(data) <- c(colnames(data)[-1],"x")
data$x <- NULL

# Renaming columns in the data to be less verbose
data <- rename(data, c("contbr_city" = "city",
                       "contbr_st" = "st",
                       "contbr_zip" = "zip",
                       "contbr_employer" = "employer",
                       "contbr_occupation" = "occupation",
                       "contb_receipt_amt" = "amount",
                       "contb_receipt_dt" = "date"))

zips <- rename(zips, "zip_code" = "zip")

# Zipcodes need cleaning because majority are postal codes, only take first 5 characters
data$zip <- substr(data$zip, 0, 5)
data$zip <- ifelse(nchar(data$zip) < 4, 
                   NA,
                   data$zip)
#subset(data$zip, nchar(data$zip) < 4)
#unique(data$zip)

mdata <- merge(data, zips[, c("zip","latitude","longitude", "state")], by = "zip")

# removing zip codes that aren't from NY state
mdata <- subset(mdata, mdata$state == "NY")
```

This report explores a dataset containing approximately 437000 Financial Contributions to the 2016 Presidential Campaigns.

I retrieved US Zip Codes from this [website](https://www.gaslampmedia.com/download-zip-code-latitude-longitude-city-state-county-csv/).
***

## Univariate Plots Section
```{r echo=FALSE, Univariate_Plots}
dim(mdata)
str(mdata)
summary(mdata)
```

```{r Long and Lat}
# Plotting the long and lat of the donations
ggplot(aes(x = longitude, y = latitude), 
       data = subset(mdata, 
                     !is.na(mdata$zip))) +
  geom_point()
```

From the plot, you can visualise the distribution of donations from NY State, however the boundaries could be more clear.

```{r Map of New York}
all_states <- map_data("state")
newyork <- subset(all_states, region %in% "new york")

# Drawing silouhette of NY with contributor data points
ggplot() +
  geom_polygon(data = newyork,
               aes(x = long, y = lat, group = group),
               colour = "grey") +
  geom_point(aes(x = longitude, y = latitude, group = NULL, size = amount), 
             data = subset(mdata, 
                     !is.na(mdata$zip)),
             colour = "green",
             alpha = 0.05) 


```


## Univariate Analysis

### What is the structure of your dataset?

### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?

***

# Bivariate Plots Section
```{r echo=FALSE, Bivariate_Plots}

```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

### What was the strongest relationship you found?


***

# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

***

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

***

# Reflection
I wanted to answer the question of whether if you come from a certain town, are you more likely to receive votes from that town.

