## Exploration of Financial Contributions for the 2016 Presidential Campaign 

#### by Sim Virdi

***

## Abstract

This report explores a dataset containing approximately 437000 Financial Contributions to the 2016 Presidential Campaigns.

I retrieved US Zip Codes from this [website](https://www.gaslampmedia.com/download-zip-code-latitude-longitude-city-state-county-csv/).

## Dataset

```{r message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
library(gridExtra)
library(RColorBrewer)
library(reshape2)
library(dplyr)
library(tidyr)
library(lubridate) # Might be needed for dates, See problem set 1
library(GGally)
library(scales)
library(memisc)
library(maps) # Needed for plotting NY State

```

```{r echo=FALSE, Load_the_Data}
# Load the Data
setwd("/Users/Sim/Desktop/Nanodegrees/Data Analyst/Task 4/Data Analysis with R/Assignment/")
data <- read.csv("NY_contributions.csv", row.names = NULL)
zips <- read.csv("zip_codes_states.csv")
parties <- read.csv("cand_gend_party.csv")

# Remove incorrect column from the dataset
colnames(data) <- c(colnames(data)[-1],"x")
data$x <- NULL

# Renaming columns in the data to be less verbose
data <- rename(data, c("contbr_city" = "city",
                       "contbr_st" = "st",
                       "contbr_zip" = "zip",
                       "contbr_employer" = "employer",
                       "contbr_occupation" = "occupation",
                       "contb_receipt_amt" = "amount",
                       "contb_receipt_dt" = "date"))

zips <- rename(zips, "zip_code" = "zip")

# Zipcodes need cleaning because majority are postal codes, only take first 5 characters
data$zip <- substr(data$zip, 0, 5)
data$zip <- ifelse(nchar(data$zip) < 4, 
                   NA,
                   data$zip)
#subset(data$zip, nchar(data$zip) < 4)
#unique(data$zip)

mdata <- merge(data, zips[, c("zip","latitude","longitude", "state")], by = "zip")
mdata <- merge(mdata, parties, by = "cand_nm")

# removing zip codes that aren't from NY state
mdata <- subset(mdata, mdata$state == "NY")

# Only interested in general and primary election from 2016
mdata <- subset(mdata, mdata$election_tp == "P2016" | mdata$election_tp == "G2016")
```

***

## Univariate Plots Section

```{r echo=FALSE, Univariate_Plots}
dim(mdata)
str(mdata)
summary(mdata)
```

### Distribution in State

In the plot below, you can visualise the distribution of donations from NY State, however it is uneasy to see both the state and city boundaries. 
```{r echo=FALSE, warning=FALSE, Long_and_Lat}
# Plotting the long and lat of the donations
ggplot(aes(x = longitude, y = latitude), 
       data = subset(mdata, 
                     !is.na(mdata$zip))) +
  geom_point()
```

To make the boundaries clearer, I used the following [tutorial](https://uchicagoconsulting.wordpress.com/tag/r-ggplot2-maps-visualization/). 

From the plot, you can see that the largest cities have the highest frequency of contributions.
```{r echo=FALSE, warning=FALSE, Map_of_New_York}
# Choosing new york state
all_states <- map_data("state")
newyork <- subset(all_states, region %in% "new york")

# Drawing silouhette of NY with contributor data points
ggplot() +
  geom_polygon(data = newyork,
               aes(x = long, y = lat, group = group)) +
  geom_jitter(aes(x = longitude, y = latitude, group = NULL), 
             data = subset(mdata, 
                     !is.na(mdata$zip)),
             alpha = 1/50,
             color = "coral1") +
  geom_point(data = subset(us.cities,
                           us.cities$country == "NY" & us.cities$pop > 100000),
             aes(x = long, y = lat, size = pop),
             colour = "green") +
  geom_text(data = subset(us.cities,
                          us.cities$country == "NY" & us.cities$pop > 100000),
            aes(x = long, y = lat, label = name),
            colour = "gold",
            fontface="bold")
```


### Amount Contributed

The next plot is used to visualise the amount of money being contributed. 

From the plot, you can see there are negative amounts being displayed due to refunds (we will explore them in further detail later).

```{r echo=FALSE, warning=FALSE, Amounts_with_refunds}
# Plot with refunds
qplot(x = amount, data = mdata) +
  scale_x_continuous(limits = c(-400, 400), breaks = seq(-400,400,50))

dim(subset(mdata, mdata$amount < 0))

# Summary with refunds.
summary(mdata$amount)

```

After removing the refunds, we see a positive skew in the distribution, with most of the money contributed between $15-$100. The median contribution is 27 with an IQR of 85. The mean is quite heavily skewed (due to v. large contributions).

```{r echo=FALSE, warning=FALSE, Amounts_without_refunds}
# Plot without refunds
qplot(x = amount, data = subset(mdata, amount >= 0)) +
  scale_x_continuous(limits = c(0, 400), breaks = seq(0,400,20))

# Summary without refunds.
summary(subset(mdata$amount, mdata$amount >= 0))
```

With the amount transformed using log10, we can see a normal distrubiton with a peak at 27.

```{r echo=FALSE, warning=FALSE, Amounts_transformed_with_log_without_refunds}
# Log transformed 
ggplot(aes(amount),
       data = subset(mdata, amount > 0)) +
  geom_histogram() +
  scale_x_continuous(trans = log10_trans(),
                     limits = c(1, 10000),
                     breaks = c(0, 10, 100, 500, 1000, 10000)) 
```

### Contributions by Gender

We are now exploring contribution amounts by gender. 

As can be seen in the histograms, both distributions have a very similar shape (both had a positive skew) with the male category having a higher frequency of contributions at $30 & $50.

From the summaries and boxplots respectively, we can see that Females have a larger mean and median contribution amount. Females also have anomalies at higher amounts.

From the frequency polygon, we get a clearer picture about the different amounts contributed to gender with females only having higher contributions at $20 and $70.
```{r echo=FALSE, warning=FALSE, Contributions_by_Gender}
# Histogram wrapped by gender
qplot(x = amount, data = subset(mdata, amount >= 0)) +
  scale_x_continuous(limits = c(0, 200), breaks = seq(0,200,20)) +
  facet_wrap(~gender)

table(mdata$gender)
by(mdata$amount, mdata$gender, summary)

# Boxplots
qplot(x = gender, y = log10(amount),
      data = subset(mdata, amount > 0),
      geom = 'boxplot')

# Frequency Plot
qplot(x = amount, y = ..count..,
      data = subset(mdata,  amount > 0), 
      xlab = 'Amount Contributed',
      ylab = 'Frequency',
      binwidth = 10, 
      geom = 'freqpoly', 
      color = gender) + 
  scale_x_continuous(lim = c(0, 200), breaks = seq(0, 200, 20)) 

```

### Contributions by Party

We are now exploring contribution amounts by Party. 

From the histogram, we can determine that the plots all have similar shapes with a positive skew. 

From the Frequency polygon, we can see that the Democratic party has significantly more contributions than the Republicans (7.37 times more) and the Green party (1585 times more).

From the summaries and boxplot, we can see that the Democrats have the lowest median and mean (27, 191 respectively) of the 3 parties. The next highest medians are the Republicans with a median of 50 although they have the highest mean of 241.5. The green party has a median of 104.0 and a mean of 201.8 although they received significantly less contributions.

```{r echo=FALSE, warning=FALSE, Contributions_by_Party}
# Histogram wrapped by party
qplot(x = amount, data = subset(mdata, amount >= 0)) +
  scale_x_continuous(limits = c(0, 200), breaks = seq(0,200,20)) +
  facet_wrap(~party, scales = 'free')

# Frequency Plot
qplot(x = amount, y = ..count..,
      data = subset(mdata,  amount > 0), 
      xlab = 'Amount Contributed',
      ylab = 'Frequency',
      binwidth = 10, 
      geom = 'freqpoly', 
      color = party) + 
  scale_x_continuous(lim = c(0, 200), breaks = seq(0, 200, 20)) 

table(mdata$party)
by(mdata$amount, mdata$party, summary)

# Boxplots
qplot(x = party, y = log10(amount),
      data = subset(mdata, amount > 0),
      geom = 'boxplot')
```

## Univariate Analysis

### What is the structure of your dataset?
There are 428646 contributions in the dataset with 23 features. 

Observations:

* Most contributions are coming from heavily populated areas such as cities.
* The median contribution is $27 and max contribution is $4904860.51.
* Contributions are spread almost equally across gender.
* Most contributions go to the Democratic Party.

### What is/are the main feature(s) of interest in your dataset?
The main features in the dataset are amount, cand_nm and party. I'd like to determine which features are the best for predicting the contribution size.

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?
Gender, occupation, election_tp, zip and time will likely influence the contribution size.

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?

I log transformed the amount of contribution for right skewed plots so it would be easier to see distributions. The distribution that I thought was most unusual was the contributions by party. I assumed that all parties would have a similar frequency of contributions however it turned out that the Democrats had almost 80% of contributions.

I adjusted column names to make them less verbose and removed rows that had data not related to the Primary or General 2016 elections or not within NY state. 

I inserted columns into the data to hold the longitude/latitude of zipcodes as well as gender and party of candidates. I did this so I could plot the contributions on a map over NY state and so that I could determine how the contributions were split between parties and gender of the candidates.

***

## Bivariate Plots Section

### Contributions and Refunds by Party

```{r echo=FALSE, Creating_table_for_party}
# Without Refunds
mdata.by_party <- subset(mdata, mdata$amount >= 0) %>%
  group_by(party) %>%
  summarise(amount_sum = sum(amount),
            amount_mean = mean(amount),
            amount_median = median(amount),
            n = n()) %>%
  arrange(party)

# With refunds
mdata.by_party_refunds <- subset(mdata, mdata$amount < 0) %>%
  group_by(party) %>%
  summarise(amount_sum = abs(sum(amount)),
            amount_mean = abs(mean(amount)),
            amount_median = abs(median(amount)),
            n = n()) %>%
  arrange(party)
```

#### Contributions

From the various barcharts and tables below, you can see that the Democratic party had the highest amount of contributions (approx $75MM) followed by the Republicans (approx $15MM) and finally the Green party had the least amount (< $1MM).

The means of the contribution however were quite different, with the Democrats having the highest mean ($278) followed by the Green party ($201) and finally the Democrats ($195). This is a little misleading however as the Green party only had 238 contributions and so anomalies were more able to skew the data.

The medians of the contribution are a more robust metric and the Green party had the highest with $104, followed by the Republicans with $50 and finally the Democrats with $27.

As discussed earlier, the Democrats had almost 7.5 times as many contributions compared to the Republicans and over 1500 times more contributions than the Green party.

This data shows that the total contribution amount was greatest for the Democrats, however the Republicans received more money per contribution. **Does this suggest richer people donate to the Republicans?**

```{r echo=FALSE, Party_Contributions}
# Making exp scale stop
getOption("scipen")
opt <- options("scipen" = 20)

p1 <- ggplot(mdata.by_party, aes(party, amount_sum, fill = party)) + 
  geom_bar(stat="identity") +
  xlab("Party") +
  ylab("Total Contributions") +
  ggtitle("2016 US Election Total Contribution by Party") +
  scale_y_continuous(breaks = seq(0, 80000000, 10000000))

p2 <- ggplot(mdata.by_party, aes(party, amount_mean, fill = party)) + 
  geom_bar(stat="identity") +
  xlab("Party") +
  ylab("Mean Contribution") +
  ggtitle("2016 US Election Mean Contribution by Party")

p3 <- ggplot(mdata.by_party, aes(party, amount_median, fill = party)) + 
  geom_bar(stat="identity") +
  xlab("Party") +
  ylab("Median Contribution") +
  ggtitle("2016 US Election Median Contribution by Party")

grid.arrange(p1,p2,p3, ncol=1)

View(mdata.by_party)
```

#### Refunds

From the various barcharts and tables below, you can see that the Republican party had the highest amount of refunds (approx $1.5MM) followed by the democrats (approx $1MM) and finally the Green party had none.

The means of the refunds followed this with the Republicans having the highest mean ($1224) followed by the Democrats ($297).

The medians of the refunds for the Republicans were highest at $484 followed by the Democrats at $100.

The number of refunds was greater for the democrats at 3383 compared to the republicans at 1248.

This data suggests that although the total refund amount was greater for the Republicans, there were more refunds for Democrats. However, as a percentage of the total contributions, the still Republicans had more people ask for refunds. 

```{r echo=FALSE, Party_Refunds}
p1 <- ggplot(mdata.by_party_refunds, aes(party, amount_sum, fill = party)) + 
  geom_bar(stat="identity") +
  xlab("Party") +
  ylab("Total Refunds") +
  ggtitle("2016 US Election Total Refunds by Party")

p2 <- ggplot(mdata.by_party_refunds, aes(party, amount_mean, fill = party)) + 
  geom_bar(stat="identity") +
  xlab("Party") +
  ylab("Mean Refund") +
  ggtitle("2016 US Election Mean Refunds by Party")

p3 <- ggplot(mdata.by_party_refunds, aes(party, amount_median, fill = party)) + 
  geom_bar(stat="identity") +
  xlab("Party") +
  ylab("Median Refund") +
  ggtitle("2016 US Election Median Refunds by Party")

grid.arrange(p1,p2,p3, ncol=1)

View(mdata.by_party_refunds)
```

### Contributions and Refunds by Candidate

```{r echo=FALSE, Creating_table_for_Candidates}
# Without Refunds
mdata.by_candidate <- subset(mdata, mdata$amount >= 0) %>%
  group_by(cand_nm) %>%
  summarise(amount_sum = sum(amount),
            amount_mean = mean(amount),
            amount_median = median(amount),
            n = n()) %>%
  arrange(cand_nm)

# With refunds
mdata.by_candidate_refunds <- subset(mdata, mdata$amount < 0) %>%
  group_by(cand_nm) %>%
  summarise(amount_sum = abs(sum(amount)),
            amount_mean = abs(mean(amount)),
            amount_median = abs(median(amount)),
            n = n()) %>%
  arrange(party)
```

#### Contributions
```{r echo=FALSE, Candidate_Contributions}
View(mdata.by_candidate)

```

#### Refunds
```{r echo=FALSE, Candidate_Refunds}

```

### Contributions and Refunds by Gender

```{r echo=FALSE, Creating_table_for_gender}
mdata.by_gender <- subset(mdata, mdata$amount >= 0) %>%
  group_by(gender) %>%
  summarise(amount_sum = sum(amount),
            amount_mean = mean(amount),
            amount_median = median(amount),
            n = n()) %>%
  arrange(gender)

# With refunds
mdata.by_gender_refunds <- subset(mdata, mdata$amount < 0) %>%
  group_by(party) %>%
  summarise(amount_sum = abs(sum(amount)),
            amount_mean = abs(mean(amount)),
            amount_median = abs(median(amount)),
            n = n()) %>%
  arrange(party)
```

#### Contributions
```{r echo=FALSE, Gender_Contributions}
View(mdata.by_gender)
```

#### Refunds
```{r echo=FALSE, Gender_Refunds}

```

### Contributions by Occupation


## Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

### What was the strongest relationship you found?


***

## Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}

```

## Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

***

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

***

# Reflection
I wanted to answer the question of whether if you come from a certain town, are you more likely to receive votes from that town.

