## Exploration of Financial Contributions for the 2016 Presidential Campaign 

#### by Sim Virdi

***

## Abstract

## Dataset
### Preperations

```{r message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
library(gridExtra)
library(RColorBrewer)
library(reshape2)
library(dplyr)
library(tidyr)
library(lubridate) # Might be needed for dates, See problem set 1
library(GGally)
library(scales)
library(memisc)
library(maps) # Needed for plotting NY State

```

```{r echo=FALSE, Load_the_Data}
# Load the Data
setwd("/Users/Sim/Desktop/Nanodegrees/Data Analyst/Task 4/Data Analysis with R/Assignment/")
data <- read.csv("NY_contributions.csv", row.names = NULL)
zips <- read.csv("zip_codes_states.csv")
parties <- read.csv("cand_gend_party.csv")

# Remove incorrect column from the dataset
colnames(data) <- c(colnames(data)[-1],"x")
data$x <- NULL

# Renaming columns in the data to be less verbose
data <- rename(data, c("contbr_city" = "city",
                       "contbr_st" = "st",
                       "contbr_zip" = "zip",
                       "contbr_employer" = "employer",
                       "contbr_occupation" = "occupation",
                       "contb_receipt_amt" = "amount",
                       "contb_receipt_dt" = "date"))

zips <- rename(zips, "zip_code" = "zip")

# Zipcodes need cleaning because majority are postal codes, only take first 5 characters
data$zip <- substr(data$zip, 0, 5)
data$zip <- ifelse(nchar(data$zip) < 4, 
                   NA,
                   data$zip)
#subset(data$zip, nchar(data$zip) < 4)
#unique(data$zip)

mdata <- merge(data, zips[, c("zip","latitude","longitude", "state")], by = "zip")
mdata <- merge(mdata, parties, by = "cand_nm")

# removing zip codes that aren't from NY state
mdata <- subset(mdata, mdata$state == "NY")

# Only interested in general and primary election from 2016
mdata <- subset(mdata, mdata$election_tp == "P2016" | mdata$election_tp == "G2016")
```

This report explores a dataset containing approximately 437000 Financial Contributions to the 2016 Presidential Campaigns.

I retrieved US Zip Codes from this [website](https://www.gaslampmedia.com/download-zip-code-latitude-longitude-city-state-county-csv/).

***

## Univariate Plots Section

```{r echo=FALSE, Univariate_Plots}
dim(mdata)
str(mdata)
summary(mdata)
```

### Distribution in State

In the plot below, you can visualise the distribution of donations from NY State, however it is uneasy to see both the state and city boundaries. 
```{r echo=FALSE, Long and Lat}
# Plotting the long and lat of the donations
ggplot(aes(x = longitude, y = latitude), 
       data = subset(mdata, 
                     !is.na(mdata$zip))) +
  geom_point()
```

To make the boundaries clearer, I used the following [tutorial on ggplot](https://uchicagoconsulting.wordpress.com/tag/r-ggplot2-maps-visualization/). 

From the plot, you can see that the larger cities have the highest frequency of contributions.
```{r echo=FALSE, Map of New York}
# Choosing new york state
all_states <- map_data("state")
newyork <- subset(all_states, region %in% "new york")

# Drawing silouhette of NY with contributor data points
ggplot() +
  geom_polygon(data = newyork,
               aes(x = long, y = lat, group = group)) +
  geom_jitter(aes(x = longitude, y = latitude, group = NULL), 
             data = subset(mdata, 
                     !is.na(mdata$zip)),
             alpha = 1/50,
             color = "coral1") +
  geom_point(data = subset(us.cities,
                           us.cities$country == "NY" & us.cities$pop > 100000),
             aes(x = long, y = lat, size = pop),
             colour = "green") +
  geom_text(data = subset(us.cities,
                          us.cities$country == "NY" & us.cities$pop > 100000),
            aes(x = long, y = lat, label = name),
            colour = "gold",
            fontface="bold") 
```


### Amount Contributed

The next plot is used to visualise the amount of money being contributed. 

From the plot, you can see there are negative amounts being displayed due to refunds (we will explore them in further detail later).

```{r echo=FALSE, Amounts with refunds}
# Plot with refunds
qplot(x = amount, data = mdata) +
  scale_x_continuous(limits = c(-400, 400), breaks = seq(-400,400,50))

dim(subset(mdata, mdata$amount < 0))

# Summary with refunds.
summary(mdata$amount)

```

After removing the refunds, we see a positive skew in the distribution, with most of the money contributed between 15 to 100 Dollars. The median contribution is 27 with an IQR of 85. The mean is quite heavily skewed (due to v. large contributions).

```{r echo=FALSE, Amounts without refunds}
# Plot without refunds
qplot(x = amount, data = subset(mdata, amount >= 0)) +
  scale_x_continuous(limits = c(0, 400), breaks = seq(0,400,20))

# Summary without refunds.
summary(subset(mdata$amount, mdata$amount >= 0))
```

With the amount transformed using log10, we can see a normal distrubiton with a peak at 27.

```{r echo=FALSE, Amounts transformed with log without refunds}

# Log transformed 
ggplot(aes(amount),
       data = subset(mdata, amount > 0)) +
  geom_histogram() +
  scale_x_continuous(trans = log10_trans(),
                     limits = c(1, 10000),
                     breaks = c(0, 10, 100, 500, 1000, 10000)) 
```

### Contributions to Male/Female Candidates

We are now plotting contribution amounts wrapped by gender. 

As can be seen, both distributions have a very similar shape with the male category having a higher frequency of contributions at $30 & $50.

From the summaries however, it can be seen that Females have a larger mean and median contribution amount.
```{r echo=FALSE, Contributions to Male/Female Candidates}
qplot(x = amount, data = subset(mdata, amount >= 0)) +
  scale_x_continuous(limits = c(0, 200), breaks = seq(0,200,20)) +
  facet_wrap(~gender)

table(mdata$gender)
by(mdata$amount, mdata$gender, summary)
```

### Contributions to Party's


## Univariate Analysis

### What is the structure of your dataset?

### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?

***

# Bivariate Plots Section
```{r echo=FALSE, Bivariate_Plots}

```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

### What was the strongest relationship you found?


***

# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

***

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

***

# Reflection
I wanted to answer the question of whether if you come from a certain town, are you more likely to receive votes from that town.

